Algorithm explanation

### 2013-09-26 article/interpolation.tex
textsc

However, by existing interpolating theorem provers, we may not be able
to obtain a simple separation.  Consider the example below.
\begin{align*}
A : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
B : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
\end{align*}
Although one may be able to obtain $x = y$ as the simplest
interpolation between $A$ and $B$, CSIsat\cite{conf/cav/BeyerZM08}
returns
\begin{align*}
I: \quad & ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
& (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
& (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
& ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
& (1 = x \vee y \leq 0)
\end{align*}
as a solution.  By the complex solution, a program verification
process may continue to abstract a program by complex predicate,
causing the verification not to terminate.

@@

\begin{alltt}
    \(\star\)
    \(\star\)
\(lbl\):
    if (\(\star\)) fail;
\end{alltt}

>>>>

However, by existing interpolating theorem provers, we may not be able
to obtain a simple separation.  Consider the example below.
\begin{align*}
A : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
B : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
\end{align*}
Although one may be able to obtain $x = y$ as the simplest
interpolation between $A$ and $B$,
\textsc{CSIsat}\cite{conf/cav/BeyerZM08} returns
\begin{align*}
I: \quad & ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
& (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
& (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
& ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
& (1 = x \vee y \leq 0)
\end{align*}
as a solution.  By the complex solution, a program verification
process may continue to abstract a program by complex predicate,
causing the verification not to terminate.

@@

\begin{alltt}
    // Do nothing
\(lbl\):
    // Non deterministic branch
    if (\(\star\)) fail;
\end{alltt}

### 2013-09-26 article/intro.tex
textsc

TODO: Mention experiments

>>>>

To confirm the effectiveness, we incorporated our algorithm into
existing software model checker \textsc{MoCHi}
\cite{conf/pldi/KobayashiSU11}.  We replaced the conventional Horn
clause solvers with our algorithms and tested some example programs to
verify the performance.  As a result, our algorithm was able to
compute a simple solution for most Horn clause solving problems.
Although we observed some test cases have experienced deterioration in
the number of verification cycles, there are some cases that are newly
verified by \textsc{MoCHi}.  Additionally, our algorithm improved a
performance in more than 70 percent of problems and achieved faster
computation than conventional method.  Eventually we have verified
certain effectiveness of our approach.

### 2013-09-26 article/jabstract.tex
textsc

\begin{jabstract}
近年のプログラム検証技術においては、自動定理証明が重要な役割を担っている。特にソフトウェア モデル検査の分野では、補間問題およびその一般化である Horn 節充足問題が注目を集めてきた。後者の問題は、循環した依存関係を持たない、未知の述語変数を複数含む Horn 節の集合の解を求める問題である。Horn 節充足問題の解、すなわち Horn 節集合を恒真とするような述語変数への代入は、プログラムの不変式の候補として用いられる。しかしながら、Horn 節充足問題の既存解法が出力する解は複雑かつ特殊すぎて不変式にならないことが多い。
上述の問題を解決するために、われわれは線形代数上の Horn 節充足問題を解く新しいアルゴリズムを提案する。われわれのアルゴリズムは、プログラム不変式が単純な式で表されるという仮定のもとで、小さな述語の生成を試みる。この目的を達成するため、このアルゴリズムはある大きさの線形不等式の雛形を未知の述語変数それぞれに対して生成し、Horn 節充足問題を線形計画法に帰着させる。そして、解が見つかるまで雛形の拡大を繰り返す。この提案手法では、合理的な時間内に比較的小さな解を構成することができる。われわれは、このアルゴリズムを実装し関数型言語向けのソフトウェア モデル検査器 MoCHi に組み込んだ。実験により、提案手法について一定の有効性を確認した。
\end{jabstract}

>>>>

\begin{jabstract}
近年のプログラム検証技術においては、自動定理証明が重要な役割を担っている。特にソフトウェア モデル検査の分野では、補間問題およびその一般化である Horn 節充足問題が注目を集めてきた。後者の問題は、循環した依存関係を持たない、未知の述語変数を複数含む Horn 節の集合の解を求める問題である。Horn 節充足問題の解、すなわち Horn 節集合を恒真とするような述語変数への代入は、プログラムの不変式の候補として用いられる。しかしながら、Horn 節充足問題の既存解法が出力する解は複雑かつ特殊すぎて不変式にならないことが多い。
上述の問題を解決するために、われわれは線形代数上の Horn 節充足問題を解く新しいアルゴリズムを提案する。われわれのアルゴリズムは、プログラム不変式が単純な式で表されるという仮定のもとで、小さな述語の生成を試みる。この目的を達成するため、このアルゴリズムはある大きさの線形不等式の雛形を未知の述語変数それぞれに対して生成し、Horn 節充足問題を線形計画法に帰着させる。そして、解が見つかるまで雛形の拡大を繰り返す。この提案手法では、合理的な時間内に比較的小さな解を構成することができる。われわれは、このアルゴリズムを実装し関数型言語向けのソフトウェア モデル検査器 \textsc{MoCHi} に組み込んだ。実験により、提案手法について一定の有効性を確認した。
\end{jabstract}

### 2013-09-26 article/related.tex
textsc

Interpolating theorem proving have played an important role in program
verification.  For instance, it is applied to invariant generation for
verification condition based program analysis, and predicate discovery
for the model checking by CEGAR with predicate abstraction.  The
former method is adopted when analyzing input programs by denotational
semantics such as Hoare logic, and synthesizing invariants from
strongest postconditions and weakest preconditions at various program
locations.  The latter one is used when model check a system with
infinite states by abstracting its data domain with predicates into a
finite system.

@@

Inspired from this method, \cite{conf/popl/HenzingerJMM04} adopted the
use of Craig interpolation for predicate discovery in CEGAR framework
with predicate abstraction, and proposed a method to derivate an
interpolant from counterexample's refutation proof derivation tree.
This work implemented the proposed method into the Blast software
model checker \cite{conf/popl/HenzingerJMS02,journals/sttt/BeyerHJM07}
and verified the effectiveness of interpolants for predicate
abstraction through an experiment of C program verification.

\paragraph{Development of interpolating theorem provers}
An interpolating prover \textsc{FOCI} \cite{website/foci} was
developed under the comprehensive work on quantifier-free interpolant
generation over the theory of linear arithmetic and uninterpreted
function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At the same
time, various applications of interpolants for model checking was
proposed \cite{conf/tacas/McMillan05}. They incorporated the algorithm
into Z3 SMT solver \cite{conf/tacas/MouraB08} and it is possible to
generate an interpolant from proofs generated by
Z3 \cite{conf/fmcad/McMillan11}.

@@

One of the most important applications of Horn clause solving is
verifying higher-order functional programs.  Interpolation is used as
a sub procedure of solving Horn clauses for dependent type
inference \cite{conf/ppdp/UnnoK09, conf/pepm/SatoUK13} in this
context. In MoCHi, the refinement types are given for values and
variables for program abstraction.  Horn clause solving is necessary
for inferring the refinment types.  \cite{conf/popl/Terauchi10}
suggested a method for dependent type inference algorithm as
well. Their algorithm is complete in the sense that it eventually
returns a solution which terminates the verification process.

>>>>

Interpolating theorem proving have played an important role in program
verification.  For instance, it is applied to invariant generation for
verification condition based program analysis, and predicate discovery
for the model checking by CEGAR with predicate abstraction.  The
former method is adopted when analyzing input programs by axiomatic
semantics such as Hoare logic, and synthesizing invariants from
strongest postconditions and weakest preconditions at various program
locations.  The latter one is used when model check a system with
infinite states by abstracting its data domain with predicates into a
finite system.

@@

Inspired from this method, \cite{conf/popl/HenzingerJMM04} adopted the
use of Craig interpolation for predicate discovery in CEGAR framework
with predicate abstraction, and proposed a method to derivate an
interpolant from counterexample's refutation proof derivation tree.
This work implemented the proposed method into the \textsc{Blast}
software model checker
\cite{conf/popl/HenzingerJMS02,journals/sttt/BeyerHJM07} and verified
the effectiveness of interpolants for predicate abstraction through an
experiment of C program verification.

\paragraph{Development of interpolating theorem provers}
An interpolating prover \textsc{FOCI} \cite{website/foci} was
developed under the comprehensive work on quantifier-free interpolant
generation over the theory of linear arithmetic and uninterpreted
function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At
the same time, various applications of interpolants for model checking
was proposed \cite{conf/tacas/McMillan05}. They incorporated the
algorithm into \textsc{Z3} SMT solver \cite{conf/tacas/MouraB08} and
it is possible to generate an interpolant from proofs generated by
\textsc{Z3} \cite{conf/fmcad/McMillan11}.

@@

One of the most important applications of Horn clause solving is
verifying higher-order functional programs.  Interpolation is used as
a sub procedure of solving Horn clauses for dependent type inference
\cite{conf/ppdp/UnnoK09, conf/pepm/SatoUK13} in this context. In
\textsc{MoCHi}, the refinement types are given for values and
variables for program abstraction.  Horn clause solving is necessary
for inferring the refinment types.  \cite{conf/popl/Terauchi10}
suggested a method for dependent type inference algorithm as well.
